### 2022-11-04 chapters/introduction.tex
update : typos

Nowadays, computers are invading our daily lives, from work to leisure, from fancy smartphones to embedded peacemakers that regulate the heartbeat of people.
As human beings, we are known to use tools to empower our bodies.
Moreover, thanks to computers, we pushed that step even further, to the point where now we are using machines to extend our brains, from equation solvers to tools to recommend to us where we should invest our money, what we should eat, and even who fits best as our partner.
One major aspect of computers that became omnipresent in our lives is the Internet, which is a network of networks that connects millions of computers all over the world.
According to Internet World Stats,\footnote{\url{https://www.internetworldstats.com/stats.htm}} the number of people connected to the Internet has increased by 4.4 billion in 2019, reaching 4.54 billion worldwide, or 59.2\% of the world population.

The Internet has evolved from a place where government researchers share information in the 60s to a means of Communication at the beginning of the century, and now it is a place where we can find almost anything we want, from information to entertainment, from social media to e-commerce.

@@

Moreover, due to the pandemic caused by COVID-19 disease, the world has been forced to adapt to a new way of living, which has been accelerated by the Internet.
The Internet has become a necessity for people to work, study~\cite{naresh2020education}, and even health consultations~\cite{liaw2021primary}.

On the other hand, as humanity, we face a major challenge, which is climate change.
The \emph{Intergovernmental Panel on Climate Change} (IPCC) has warned that the world has only 12 years to limit the global temperature rise to 1.5\degree C and that the world has to reach net-zero emissions by 2050 to avoid the worst effects of climate change.
The IPCC has also warned that the world has to reduce its emissions by 45\% by 2030 to reach the 1.5\degree C target~\cite{portner2022climate}.

To survive, we have come up with three solutions.
The first one includes finding a new planet that we can populate and live on,\footnote{\url{https://en.wikipedia.org/wiki/Interstellar_(film)}} which is known as the Planetary Migration~\cite{mapstone2022cyanobacteria}.
Meanwhile, the second solution is to provide new sources of energy, such as nuclear energy, wind energy, and even fusion energy~\cite{gross1984fusion}.
The third solution is to reduce our emissions, which is the main focus of this thesis.

@@

With this increase in the number of data centers comes an increase in energy consumption, which is a major problem for the environment.
in 2018, data centers consumed around 205 terawatt-hours (TWh)~\cite{schneider2021world}, which is equivalent to the energy consumption of 1\% of the total world's electricity.
This ratio increased up to 1.5 \% in 2020 according to the Journal of Science~\cite{mytton2021data}.
In Figure~\ref{fig:data_centers_power_distribution}, one can see that 40\% of the energy consumed by data centers is used for cooling, while another 40\% is used by the servers themselves.
Therefore, optimizing these two aspects can have a major impact on the energy consumption of data centers.

@@

Researchers are trying to reduce the energy consumption of data centers through different angles.
Some of the works are focused on the hardware side, such as using new hardware architectures that are more energy-friendly, such as the use of GPUs instead o ARM processors instead of CPUs~\cite{aroca2012towards}.
Others are trying to optimize the cooling system, this can be achieved by using more efficient cooling systems, putting data centers in cold locations or under water~\cite{simon2018project}, or even using the waste heat for other purposes, such as heating buildings~\cite{bouzel2021distributed,cao2021carbon}.

A third approach is to optimize the software, by making software more energy-efficient.
In this thesis, we focus on this approach, and we try to optimize the software by reducing the number of computations that are done by the software.

The best way to do so is to formulate a theory behind the energy consumption of algorithms, such as the complexity and the o notation.
Unfortunately, this is not possible in the current state of the art.
Due to the lack of knowledge about the energy consumption of the algorithms, and the strong correlation between this consumption and the hardware configuration.
Unlike algorithm optimization in the field of performance, which is agnostic toward the platform, the energy consumption of the algorithms is dependent on the execution environment.
Therefore, for the moment, we start by formulating some hypotheses and exploring them using empirical analysis.
Figure~\ref{fig:thesis_position} highlights this thesis's position on the sustainability of ICT\@; while this thesis only addresses a small portion of ICT's energy usage, we feel it is a step in the right direction for additional solutions to mature to preserve humanity.

@@

The final strategy consists of \emph{deliberately reducing the performance of the software to reduce its energy consumption}.
The goal of this method is to reduce energy usage by sacrificing execution time.
This strategy is useful while dealing with services since the task has an unlimited execution time.
The second part is to use some. opportunistic schedulers, which will stop the software if it is estimated to spend more energy.
This strategy is discussed first in Chapter~\ref{chapter:porgramming_langauges}, then we continue with this strategy in the perspectives section.

@@

          First, we discuss the reproducibility challenge and how to overcome it with the usage of containers, then we push this approach further by proposing and new protocol to make tests not only reproducible but able to be extended with other features to overcome the rapid pace of software evolution.
          The second part of this chapter discusses the aspect of accuracy when it comes to energy consumption measurements.

@@

          For each use case, we compare the energy consumption of several approaches and provide guidelines on how to optimize the energy consumption of python programs in these use cases.
          Finally, we provide a non-intrusive way to optimize energy consumption without altering the code of the program.

@@

          We compare the energy consumption of $12$ benchmarks using $52$ implementation, each benchmark is dedicated to a typical use case.

@@

          This analysis aims to look at how each technology uses energy in different kinds of web situations.
    \item Finally, we conclude our work in Chapter~\ref{chapter:conclusion} by summarizing the work done in this thesis and discussing the future work that can be done to further reduce the energy consumption of software.

>>>>

Nowadays, computers are invading our daily lives, from work to leisure, from fancy smartphones to embedded peacemakers that regulate the heartbeat of people.
As human beings, we are known to use tools to empower our bodies.
Moreover, thanks to computers, we pushed that step even further, to the point where now we are using machines to extend our brains, from equation solvers to tools to recommend to us where we should invest our money, what we should eat, and even who fits best as our partner.
One significant aspect pushing computers to become omnipresent in our lives is the Internet, a network connecting millions of computers worldwide.
According to Internet World Stats,\footnote{\url{https://www.internetworldstats.com/stats.htm}} the number of people connected to the Internet has increased by 4.4 billion in 2019, reaching 4.54 billion worldwide, or 59.2\% of the world population.

The Internet has evolved from a place where government researchers share information in the 60s to Communication support at the beginning of the century. Now it is where we can find almost anything we want, from information to entertainment, social media to e-commerce.

@@

Moreover, due to the pandemic caused by COVID-19 disease, the world has been forced to adopt a new living style, in which the Internet has become a necessity for people to work, study~\cite{naresh2020education}, and even health consultations~\cite{liaw2021primary}.

On the other hand, as humanity, we face a significant challenge, which is climate change.
The \emph{Intergovernmental Panel on Climate Change} (IPCC) has warned that the world has only 12 years to limit the global temperature rise to 1.5\degree C and that the world has to reach net-zero emissions by 2050 to avoid the worst effects of climate change.
The IPCC has also warned that the world has to reduce its emissions by 45\% by 2030 to reach the 1.5\degree C target~\cite{portner2022climate}.

To survive, we have come up with three solutions.
The first one includes finding a new planet that we can populate and live on,\footnote{\url{https://en.wikipedia.org/wiki/Interstellar_(film)}} which is known as the Planetary Migration~\cite{mapstone2022cyanobacteria}.
Meanwhile, the second solution is to provide new energy sources, such as nuclear energy, wind energy, and even fusion energy~\cite{gross1984fusion}.
The third solution is to reduce our emissions, which is the main focus of this thesis.

@@

With this increase in the number of data centers comes an increase in energy consumption, which is a significant environmental problem.
In 2018, data centers consumed around 205 terawatt-hours (TWh)~\cite{schneider2021world}, equivalent to the energy consumption of 1\% of the entire world's electricity.
This ratio increased up to 1.5\% in 2020 according to the Journal of Science~\cite{mytton2021data}.
In Figure~\ref{fig:data_centers_power_distribution}, one can see that 40\% of the energy consumed by data centers is used for cooling, while another 40\% is used by the servers themselves.
Therefore, optimizing these two aspects can significantly impact the energy consumption of data centers.

@@

Researchers are trying to reduce the energy consumption of data centers through different angles.
Some of the works are focused on the hardware side, such as using new hardware architectures that are more energy-friendly, such as using GPUs instead o ARM processors instead of CPUs~\cite{aroca2012towards}.
Others are trying to optimize the cooling system; this can be achieved by using more efficient cooling systems, putting data centers in cold locations or under water~\cite{simon2018project}, or even using the waste heat for other purposes, such as heating buildings~\cite{bouzel2021distributed,cao2021carbon}.

A third approach is to optimize the software by making software more energy-efficient.
In this thesis, we focus on this approach and try to optimize the software by reducing the number of computations done by the software.

The best way to do so is to formulate a theory behind the energy consumption of algorithms, such as the complexity and the o notation.
Unfortunately, this is not possible in the current state of the art.
Due to the lack of knowledge about the energy consumption of the algorithms and the strong correlation between this consumption and the hardware configuration.
Unlike algorithm optimization in the field of performance, which is agnostic toward the platform, the energy consumption of the algorithms is dependent on the execution environment.
Therefore, for the moment, we start by formulating some hypotheses and exploring them using empirical analysis.
Figure~\ref{fig:thesis_position} highlights this thesis's position on the sustainability of ICT\@; while this thesis only addresses a small portion of ICT's energy usage, we feel it is a step in the right direction for additional solutions to mature to preserve humanity.

@@

The final strategy consists of \emph{deliberately reducing the performance of the software to reduce its energy consumption}.
The goal of this method is to reduce energy usage by sacrificing execution time.
This strategy is helpful while dealing with services since the task has an unlimited execution time.
The second part is to use some opportunistic schedulers, which will stop the software if it is estimated to spend more energy.
This strategy is discussed first in Chapter~\ref{chapter:porgramming_langauges}, then we continue this strategy in the perspectives section.

@@

            First, we discuss the reproducibility challenge and how to overcome it with the usage of containers; then we push this approach further by proposing and new protocol to make tests not only reproducible but able to be extended with other features to overcome the rapid pace of software evolution.
            The second part of this chapter discusses the aspect of accuracy regarding energy consumption measurements.

@@

            For each use case, we compare the energy consumption of several approaches and provide guidelines on optimizing the energy consumption of python programs in these use cases.
            Finally, we provide a non-intrusive way to optimize energy consumption without altering the program's code.

@@

            We compare the energy consumption of $12$ benchmarks using $52$ implementation, and each benchmark is dedicated to a typical use case.

@@

            This analysis aims to look at how each technology uses energy in different web situations.
      \item Finally, we conclude our work in Chapter~\ref{chapter:conclusion} by summarizing the work done in this thesis and discussing the future work that can be done to reduce the energy consumption of software further.
